Transformation of forecasts for evaluating predictive performance in an epidemiological context

### 2023-01-24 main.tex
update citation
Updates from Overleaf

Forecast evaluation plays an essential role in the development cycle of predictive epidemic models and can inform their use for public health decision-making. Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which are both measures of the absolute distance between the forecast distribution and the observation. They are commonly applied directly to predicted and observed incidence counts, but it can be questioned whether this yields the most meaningful results given the exponential nature of epidemic processes and the several orders of magnitude that observed values can span over space and time. In this paper, we argue that log transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results. We motivate the procedure threefold using the CRPS on log-transformed counts as an example: Firstly, it can be interpreted as a probabilistic version of a relative error. Secondly, it reflects how well models predicted the time-varying epidemic growth rate. And lastly, using arguments on variance-stabilizing transformations, it can be shown that under the assumption of a quadratic mean-variance relationship, the logarithmic transformation leads to expected CRPS values which are independent of the order of magnitude of the predicted quantity. Applying the log transformation to data and forecasts from the European COVID-19 Forecast Hub, we find that it changes model rankings regardless of stratification by forecast date, location or target types. Situations in which models missed the beginning of upward swings are more strongly emphasized while failing to predict a downturn following a peak is less severely penalized. We conclude that appropriate transformations, of which the natural logarithm is only one particularly attractive option, should be considered when assessing the performance of different models in the context of infectious disease incidence. 
\end{abstract}

>>>>

Forecast evaluation plays an essential role in the development cycle of predictive epidemic models and can inform their use for public health decision-making. Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which are both measures of the absolute distance between the forecast distribution and the observation. They are commonly applied directly to predicted and observed incidence counts, but it can be questioned whether this yields the most meaningful results given the exponential nature of epidemic processes and the several orders of magnitude that observed values can span over space and time. In this paper, we argue that transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results. We motivate the procedure threefold using the CRPS on log-transformed counts as an example: Firstly, it can be interpreted as a probabilistic version of a relative error. Secondly, it reflects how well models predicted the time-varying epidemic growth rate. And lastly, using arguments on variance-stabilizing transformations, it can be shown that under the assumption of a quadratic mean-variance relationship, the logarithmic transformation leads to expected CRPS values which are independent of the order of magnitude of the predicted quantity. Applying the log transformation to data and forecasts from the European COVID-19 Forecast Hub, we find that it changes model rankings regardless of stratification by forecast date, location or target types. Situations in which models missed the beginning of upward swings are more strongly emphasized while failing to predict a downturn following a peak is less severely penalized. We conclude that appropriate transformations, of which the natural logarithm is only one particularly attractive option, should be considered when assessing the performance of different models in the context of infectious disease incidence. 
\end{abstract}
